\documentclass[a4paper,11pt]{article}
\usepackage[top=1in,bottom=1in,left=1.25in,right=1.25in]{geometry}
\linespread{1.25}

\title{SoLISP Design}
\author{Yuheng Kuang (kuangyuheng@gmail.com)}

\begin{document}
\maketitle

\section{Introduction}

\subsection{Simple facts}

\begin{enumerate}
\item SoLISP is a LISP variants. It uses S-Expression and has common syntax like \verb|(+ 1 value)|
\item SoLISP aims to map Python language feature to S-Expression, instead of implementing yet another dynamic language.
\item SoLISP shares same data and execution with Python. You can use Python libraries you love in SoLISP directly.
\item SoLISP compiles into plain human readable Python code.
\item Beyond Python, SoLISP provide features like Macro system and Pattern matching (Proc model).
\end{enumerate}

\subsection{Why SoLISP?}

\section{A taste of SoLISP}

\subsection{Hello World}

Say hello world using sys.stdout.write. It represent the LISP natual of SoLISP and how it integerate with Python.

\begin{verbatim}
(import sys)
(def (hello :name "World")
  (sys -> stdout write ! (% "Hello %s!\n" name)))
(hello) ; => "Hello World!\n"
(hello "Yuheng") ; => "Hello Yuheng!\n"
\end{verbatim}

The code will be compiled into plain Python code:

\begin{verbatim}
import sys
def hello(name = 'World'):
  sys.stdout.write('Hello %s!\n', name)
hello()
hello('Yuheng')
\end{verbatim}

\subsection{Everything is expression}

As in LISP, Everything in SoLISP is expression.

\begin{verbatim}
(def (random_add :large False)
  (+ (if large 10000 else 1) 1))

(print (random_add)) ; => 1
(print (random-add True)) ; => 10001
\end{verbatim}


\subsection{Powerful for}

A simple list comprehension example.

\begin{verbatim}
(print (for x <- [1 4 2 3]
          (if (> x 2) (cont)) (* x x)))
; => [1 4]
\end{verbatim}

A nested example.

\begin{verbatim}
(print
 (for x <- ["a" "b"]
    (emit* (for y <- [1 2] [x y]))) (cont)))
; => [["a" 1] ["a" 2] ["b" 1] ["b" 2)]
\end{verbatim}


\subsection{Restricted loop}

Looping in SoLISP is more restricted. You specify init value of the loop, then use \verb|(cont next_value)| to loop, the loop will break if you do nothing.

\begin{verbatim}
(print
 (loop total <- 1
       (if (< total 1000) (cont (* total 2)))))
; => 1024
\end{verbatim}

\subsection{Pattern matching}

The test\_proc is a one-argument function, that returns input when input is int and larger than 0, else raise an \verb|MatchException|.

\begin{verbatim}
(= test_proc (# x :int ?(> x 0)))
(test_proc 1)  ; => 1
(test_proc "1") ; => raise
(test_proc 0) ; => raise
\end{verbatim}

Pattern matching can be used in assignment.

\begin{verbatim}
(= [0 x . remain] src)
\end{verbatim}

SoLISP will check the value of src's first element (equals 0?), then assign second element to x and remaining elements to remain. It's called structual matching.

\subsection{Proc: try-match-do model}

Pattern matching is only a very limited case of the powerful Proc engine.

\begin{verbatim}
(= commander
 (# (str.split ["add" (int x) (int y)]) => (+ x y)
  # (str.split ["dec" (int x) (int y)]) => (- x y)
  # => -1))

(commander "dec 10 2") => 8
(commander "add 1 a") => -1
(commander "blahblahblah") => -1
\end{verbatim}

It also shows how SoLISP can match virtually everything. The first two section of the proc split input string into list, then matching it into a list pattern. It's called ``Extractor'', we can use extractor to convert any object into basic data structure and match them.

\subsection{Proc + Looping}

Proc can be used in ``\verb|for|'' and ``\verb|loop|''.

Concise version of list comprehension. 

\begin{verbatim}
(for (# x ?(<= x 2) => (* x x)) <- [4 1 3 2])  ; => [1 4]
\end{verbatim}

\begin{verbatim}
(loop (# x ?(< x 1000)) <- 1 (cont (* x 2)))
\end{verbatim}

\subsection{Macros}

We use ``match'' expression everywhere.

\begin{verbatim}
(match x
 # ?(< _ 0) => "<"
 # ?(== _ 0) => "=="
 # => ">")
\end{verbatim}

This expression is entirely implemented in SoLISP, by defining a macro.

\begin{verbatim}
(= macro_match (# ['match value . proc] => `('= proc value)))
\end{verbatim}

It translates \verb|(match value . proc)| into \verb|(= proc value)|, then assignment primitive do the job.

\subsection{Full example}

Look into \verb|src/web/| directory in Chord package for a full example.

\section{S-Expr and basic data types}

\section{Basic language features}

\subsection{Simple expression}

\subsection{Control structure}

\section{Proc model}

\subsection{Introduction to pattern matching}

\subsection{Try - match - do model}

\subsection{Looping with pattern matching}

\section{Customize SoLISP}

\end{document}
